---
title: "Data exploration"
format: 
  html:
    echo: false
    reference-location: margin
    embed-resources: true # Self contained document.
    fig.width: 10
    fig.height: 8
toc: true
editor: visual
---

```{r}
#| echo: false
#| warning: false

library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(skimr)
library(lubridate)

responses <- read.csv("../data/raw/cyber_expertise_diversity_survey_responses.csv") |> 
  mutate(recorded_date = as.Date(recorded_date))
scenarios <- read.csv("../data/raw/cyber_expertise_diversity_survey_scenarios.csv")

```

### Responses

The survey was conducted bewteen `r min(responses$recorded_date)` and `r max(responses$recorded_date)` and received responses from `r nrow(responses)` individuals and describing `r nrow(scenarios)` scenarios.

```{r}
#| column: page

skim(responses)
```

### Scenarios

```{r}

#| column: page

skim(scenarios)
```

## Alignment metric

Explain how the metric was calculated

Given a set of **test values** $X:[x_0, x_1, ... x_n ]$ and a corresponding set of **reference values** $Y:[y_0, y_1, ... y_n ]$

$$
Asc_{(x,y)} = \frac{\sum I(X)}{\sum Y}
$$

Where the indicator function $I(X)$ is defined as follows:

-   $I(X) = x_m$ when $x_m <= y_m$
-   $I(X) = y_m$ when $x_m > y_m$

All the indicator function does is say that a person can’t get ‘credit’ for overshooting the reference values (which could offset their deficit in other areas). So it takes the reference value as a ceiling in each tooth of the test comb.

Computations for the alignment metric can be audited at the file `alignment_metric.R`

### Expertise Alignment

In this case, we are interested in how well

### Scenarios alignment

We are interested in how well the specialisms align with the scenarios. To do so, we will be calculating an *scenario alignment index (Asc)*, where the test values correspond to , and the reference values correspond to

## Megacomb

### Expertise comb

```{r}
#| column: page
#| fig-height: 15
#| fig-width: 13
#| warning: false
df <- responses |> 
  select(gender, starts_with("expertise_comb_")) |> 
  pivot_longer(!gender) |> 
  mutate(name = str_remove(name, "expertise_comb_"))

ggplot(df, aes(y = name, x = value, fill = gender)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="top")
```

### Scenarios and gender

A first draft for the megacomb:

```{r}
#| column: page
#| fig-height: 15
#| fig-width: 13
#| warning: false
scenario_risk_gender <- scenarios |> 
  select(gender, starts_with("scenario_expertise_portfolio_")) |> 
  pivot_longer(!gender) |> 
  mutate(name = str_remove(name, "scenario_expertise_portfolio_"))

ggplot(scenario_risk_gender,aes(y = name, x = value, fill = gender)) +
  geom_boxplot() +
  theme_minimal() +
  theme(legend.position="top")

```
