---
title: "Interdisciplinarity"
---

```{r}
source("../scripts/setup.R")

scenarios <- read.csv("../data/raw/cyber_expertise_diversity_survey_scenarios.csv")

alignment_scenarios_long <- read.csv("../data/processed/alignment_scenarios_long.csv")
max_alignment_scenarios_1_permutations <- read.csv("../data/processed/max_alignment_scenarios_1_permutations.csv")
max_alignment_scenarios_2_permutations <- read.csv("../data/processed/max_alignment_scenarios_2_permutations.csv")
max_alignment_scenarios_3_permutations <- read.csv("../data/processed/max_alignment_scenarios_3_permutations.csv")
max_alignment_scenarios_4_permutations <- read.csv("../data/processed/max_alignment_scenarios_4_permutations.csv")
max_alignment_scenarios_5_permutations <- read.csv("../data/processed/max_alignment_scenarios_5_permutations.csv")

max_alignment_scenarios <- max_alignment_scenarios_1_permutations |>
  bind_rows(max_alignment_scenarios_2_permutations) |> 
  bind_rows(max_alignment_scenarios_3_permutations) |> 
  bind_rows(max_alignment_scenarios_4_permutations) |> 
  bind_rows(max_alignment_scenarios_5_permutations) 

responses <- read.csv('../data/raw/cyber_expertise_diversity_survey_responses.csv') |> 
  mutate(recorded_date = dmy_hm(recorded_date))

# GGplot defaults
theme_set(theme_minimal())

```

## Interdisciplinary cyber security

In this section we were interested in the intrinsic interdisciplinary of cyber security. Understanding the value of diversity of expertise in cyber security requires a view of the problems that arise in cyber security, and the kinds of combinations of expertise that they require. For organisations, the utility of classifications of individuals into standardised specialisms depends on the extent to which it is possible to identify a need for particular kinds of individuals. Understanding how needs are aligned with specialisms means understanding the kinds of profiles of expertise that are required to address specific problems. The Cyber Expertise Diversity Survey elicited such profiles from participants, for a range of scenarios. While the dataset is limited, the presence of qualitative descriptions of these scenarios enables us to explore the ways in which scenarios map onto specialisms, in particular:

-   Which scenarios map straightforwardly onto one specialism

-   Which scenarios require combinations of specialisms

-   Which scenarios require significant expertise that is not accounted for by the specialisms

## Data analysis

### Looking at specialisms alignment with the scenarios

We calculated the alignment of the UKCSC specialisms with the scenarios our respondents told us about. We were interested in how alignment scores would increase if we allowed permutations of two or more specialisms. @fig-alignment-scenarios-overview shows that, overall, the alignment index increases as we increase the number of permutations (this is, these scenarios are more interdisciplinary). While this could be expected, this is not the case for every scenario, as shown in @fig-alignment-scenarios, which provides a more nuanced view by displaying how the alignment score changes for every scenario.

```{r}
#| label: fig-alignment-scenarios-overview
#| fig-cap: "Alignment indices per number of permutations"
max_alignment_scenarios |> 
  mutate(
    n_permutations_txt = case_when(
      n_permutations == 1 ~ "Single specialism",
      .default = paste(n_permutations, "Permutations")),
    n_permutations_txt = as.factor(n_permutations_txt),
    n_permutations_txt = fct_relevel(n_permutations_txt, "Single specialism")) |> 
  ggplot(aes(x = alignment_index * 100)) +
  geom_boxplot(outliers = FALSE) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) +
  facet_wrap(~n_permutations_txt, ncol = 1) +
  labs(x = "Alignment index") +
  theme(
    panel.grid.major.y = element_line(size=0.5, colour = "white"),
    panel.grid.minor.y = element_line(size=0.1, colour = "white"),
    axis.title.x=element_blank(),
    axis.text.y=element_blank()
  )
```

=

Scenarios with a steep gradient show indicate significantly better alignment could be gained through teams of mixed specialists.

**In the viz below, can there be more space to display the scenario names, perhaps allowing two lines of text for each? It doesn't matter if some are cut off, but a few more words would help.**

_I have incereased the text length here too 120 characters, but I think it might be better if we manually change some of them_

_Version 1_

```{r}
#| label: fig-alignment-scenarios
#| fig-cap: "Alignment indices per number of permutations and scenario"
#| column: page
#| fig-width: 14
#| fig-height: 25
#| fig-align: "center"

scenario_names <- scenarios |> 
  select(scenario_name) |> 
  mutate(id = row_number())

max_alignment_scenarios |> 
  select(id, n_permutations, alignment_index) |> 
  left_join(scenario_names, by = join_by(id)) |>
  mutate(scenario_name = paste(id,"-", scenario_name),
         scenario_name = case_when(
           nchar(scenario_name) > 120 ~ paste0(str_sub(scenario_name, end = 120), "..."),
           .default = scenario_name),
         # scenario_name = str_wrap(scenario_name, width = 30),
         scenario_name = as.factor(scenario_name),
         scenario_name = fct_reorder(scenario_name, id)) |> 
  pivot_longer(alignment_index) |> 
  ggplot(aes(x = n_permutations, y = value * 100)) + 
  geom_line(color = "red", size = 1) +
  geom_point(color = "red")+ 
  # facet_wrap( ~id, ncol = 5, 
  #             labeller = as_labeller(
  #               function(string, suffix = scenario_names$scenario_name) paste(string, suffix)
  #             )) +   
  facet_wrap(~scenario_name, ncol = 5,
            labeller = labeller(scenario_name = label_wrap_gen(width = 30))) + 
  labs(title = "Alignment indices per number of permutations and scenario",
       y = "Scenario Alignment index",
       x = "Number of specialisms permutations") +
  theme(
    panel.grid.major.x = element_line(size=0.5, colour = "white"),
    panel.grid.minor.x = element_line(size=0.1, colour = "white")
  )

```

_Version 2_

```{r}
#| label: fig-alignment-scenarios_v2
#| fig-cap: "Alignment indices per number of permutations and scenario"
#| column: page
#| fig-align: "center"
scenario_names <- scenarios |> 
  select(scenario_name) |> 
  mutate(id = row_number())

all_align <- max_alignment_scenarios |> 
  select(id, n_permutations, alignment_index) |> 
  left_join(scenario_names, by = join_by(id)) |>
  mutate(scenario_name = paste(id,"-", scenario_name),
         scenario_name = case_when(
           nchar(scenario_name) > 25 ~ paste0(str_sub(scenario_name, end = 25), "..."),
           .default = scenario_name),
         scenario_name = as.factor(scenario_name),
         scenario_name = fct_reorder(scenario_name, id)) |> 
  pivot_longer(alignment_index)

all_align <- all_align |> mutate(color = 'all')

all_mean <- all_align |> group_by(n_permutations) |> summarise(value = mean(value, na.rm=TRUE))
all_mean <- all_mean |> mutate(id = 0, scenario_name = 'all', name = 'alignment_index_all', color = 'mean')

all_align <- rbind(all_align, all_mean)

all_align <- all_align %>% filter(value > 0)
# 
# ggplot(all_align, aes(x = n_permutations, y = value, group = scenario_name, color = color)) + 
#   geom_line(size = 1, alpha = 0.2) +
#   geom_point(alpha = 0.2)


ggplot(all_align, aes(x = n_permutations, y = value * 100, group = scenario_name)) + 
  geom_line(data = all_align %>% filter(color == 'all'),
    linewidth = 1, alpha = 0.2,
    color = '#F3766E') +
  geom_point(data = all_align %>% filter(color == 'all'),
             alpha = 0.2,
             color = '#F3766E') +
  geom_line(data = all_align %>% filter(color != 'all'),
    linewidth = 1, alpha = 1,
    color = '#1CBDC2') +
  geom_point(data = all_align %>% filter(color != 'all'),
             alpha = 1,
             color = '#1CBDC2',
             size = 4) + 
    labs(title = "Alignment indices per number of permutations and scenario",
       y = "Scenario Alignment index",
       x = "Number of specialisms permutations") +
  theme_classic()

```

```{r}
#| fig-width: 14

sc_tab <- data.frame(`Column 1` = c(scenario_names$scenario_name[1:22]),
                     `Column 2` = c(scenario_names$scenario_name[23:44]),
                     `Column 3` = c(scenario_names$scenario_name[45:65], NA))

opts <- options(knitr.kable.NA = "")

sc_tab|>
  kable()
```

### Meta view of alignment vs interdisciplinarity

_New Graph_
_This is one of the graphs I made for the poster, I have included a basis version here, but can refine further if you would like to keep it_

```{r}
#| label: fig-alignment-persona-vs-scenarios
#| fig-cap: "Personal vs Scenarios alignment"
#| column: page
#| fig-width: 14
#| fig-align: "center"

response_exp <- responses[8:33]
response_exp <- response_exp |> tibble::rowid_to_column()             

response_exp <- pivot_longer(response_exp,
                            cols = names(response_exp)[2:27],
                            values_to = 'values')

response_exp <- response_exp |>
  mutate(name = str_remove(name, 'expertise_comb_'))
response_exp <- response_exp |>
  mutate(comb_type = 'personal')


response_sce <- responses[41:66]
response_sce <- response_sce |> tibble::rowid_to_column()             

response_sce <- pivot_longer(response_sce,
                            cols = names(response_sce)[2:27],
                            values_to = 'values')
response_sce <- response_sce |>
  mutate(name = str_remove(name, 'scenario_expertise_portfolio_'))
response_sce <- response_sce |>
  mutate(comb_type = 'scenario')

response_long <- rbind(response_exp, response_sce)

colnames(response_sce) <- c('rowid_sce', 'name_sce', 'values_sce', 'comb_type_sce')
colnames(response_exp) <- c('rowid_exp', 'name_exp', 'values_exp', 'comb_type_exp')

response_wide <- cbind(response_exp, response_sce)

response_wide <- response_wide %>% mutate(Color = ifelse(values_exp >= values_sce, "green", "red"))
response_wide <- response_wide %>% mutate(color_n = ifelse(values_exp >= values_sce, -100, 100))

response_wide <- response_wide %>% mutate(color_n = color_n + values_exp + values_sce)


response_wide <- response_wide |>
                    group_by(name_exp) |>
                    mutate(within_value_order = order(order(color_n,  decreasing = TRUE)))

ggplot(response_wide) +
  geom_segment(aes(y=values_exp, yend=values_sce, x=within_value_order, color = Color,), linewidth = 0.2) +
  #geom_point( aes(x=x, y=value1), color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  #geom_point( aes(x=x, y=value2), color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  #theme_minimal() +
  theme_classic() +
theme(
  strip.background = element_blank(),
  #strip.text.x = element_blank(),
  #strip.text.y.left = element_text(angle = 90),
  #panel.spacing = unit(3, "pt"),
) +
  facet_wrap(~name_sce, ncol = 26, 
             strip.position = 'bottom')
```

### Qualitative analysis of the scenario descriptions

#### Lowest alignment scenarios

Scenarios that have lower overall alignment tend to be scenarios where respondents set the combs particularly 'full'. For instance, for *Scenario 3: global mass use of malware*, the participant set all sliders to 100 (or nearly). *Scenario 13: investigation of a cyber incident*: apart from a few (data science, hardware, cyber physical and cryptography), this was deemed to require a very 'full' comb. Similarly, *Scenario 28: Develop effective and real world governance to protect assets while facilitating and enabling business goals and objectives* had a very 'full' comb.

#### Most interdisciplinary scenarios

Scenarios that had a steep gradient may be interpreted as particularly interdisciplinary. *Scenario 7: Protecting Legacy equipment against emerging AI/PQC Threat*, for instance combined significant elements of hardware security, cryptography, network security, authentication, authorisation and accountability, and distributed systems security. This hit a combination of expertise that is not directly associated within existing specialisms. No UKCSC specialism, for instance, is identified with authentication, authorisation and accountability *and* hardware security. This scenario is about protecting legacy equipment in a CNI context, and while this is a specific context it is a widely recognised problem scenario.

Scenario 27: incident response was similarly intriguing for its interdisciplinarity, given that there is a specialism called 'incident response'. The respondent had specified the problem scenario including the requirement for very high expertise in authorisation, authentication and accountability, and quite high in operating systems and virtualization security. Neither of these are taken to be of even 'wider' relevance to incident response or digital forensics specialisms.

Both scenarios 7 and 27 provide anecdotal evidence that the UKCSC CyBOK mapping may itself require further thought. Further research into expertise diversity could help test this hypothesis and improve the mapping.

Scenarios 61 and 62 were more specialist problems

::: {.callout-note appearance="minimal"}
Got pulled from a job testing building management/HVAC systems to taking lead on assessing encrypted VoIP systems. Had to give a crash course on OT testing to more junior colleagues before switching assignments (61)
:::

Scenario 61 involved a wide spread of expertise, while scenario 62 required high expertise in a relatively high number of areas.

::: {.callout-note appearance="minimal"}
Requirement to build out a cyclical, threat detection and engineering process which required:

-   Understanding of Primary Threat Adversaries and identification of their TTPs.

-   Threat Hunting capability for aforementioned TTPs (and other Indicators of Compromise).

-   Detection Engineering capability (for creation of signals, signatures and detections, specific to the aforementioned TTPs), delivering these into production.

-   Event Triage and Investigation to Incident esclation including containment, eradication and recovery.

-   Feedback mechansim of observed incidents, back into Intelligence teams to close the loop. (Detection Lifecycle Management).

    \(62\)
:::

These scenarios are the more 'naturally' interdisciplinary cases, where problems need to be addressed that are not neatly contained within standard specialism areas, and will either require individuals with unusual skillsets or else mixed teams.

#### Non-CyBOK expertise

Finally, some scenarios were specified in ways that heavily emphasised the need for non-CyBOK areas of expertise. In these cases, the alignment scores are deceptive, because we cannot calculate which specialists would contribute these skills. Interestingly, one of these cases was *58 - Incident Response is a business requirement, not just for the IT/Cyber team(s).*

::: {.callout-note appearance="minimal"}
Often people think that cyber maturity, resilience, and response is just for the cyber or IT teams to do. Promoting that it there are multiple groups that will be affected and impacted by a successful compromise and making them aware and appreciate their role to play and to consider what their individual playbook would look like is a difficult but important task to achieve. Evidence based metrics to promote the risk and then cyber resilience exercises involving key stakeholders that then promote wider engagement down the line is a necessary tool to have in your armour.
:::

Incident response is supposed to be captured within the existing specialisms, but this respondent specifies the problem in a way that emphasises expertise that is not considered within the Careers Framework. In particular, the organisational, human factors and culture aspects are deemed particularly important. Again anecdotal, examples like this are suggestive of how the process of professionalisation may be refined, understanding where the hidden areas of expertise are that are needed. Further research may help refinement of definitions, to identify areas in which the expertise needed to tackle problems is changing, and to identify areas in which there are diverse, competing, problem definitions. In the latter case, it would be important for policymakers to ensure that problem definitions enshrined in standard classifications do not suppress innovation by imposing a standard vision of what the solution should look like.

## Recommendations

We recommend that the UK government supports further research into cyber security expertise diversity, including looking at cyber security problem scenarios, to identify gaps in existing professional classifications and to help validate the Career Framework CyBOK mapping.
