---
title: "The survey"
format: html
---

## Methodology

```{r}
#| echo: false
#| warning: false

library(ggplot2)
library(gridExtra)
library(ggbeeswarm)
library(dplyr)
library(tidyr)
library(stringr)
library(skimr)
library(lubridate)
library(knitr)

#responses <- read.csv("../data/raw/cyber_expertise_diversity_survey_responses.csv") |> 
#  mutate(recorded_date = dmy_hm(recorded_date))

# sorry for changing this, for some reason I just could not get relative paths working
responses <- read.csv('/Users/u2272744/Downloads/cyberexpertisediversity_survey_data-main/data/cyber_expertise_diversity_survey_responses.csv') |> 
  mutate(recorded_date = dmy_hm(recorded_date))


scenarios <- read.csv("/Users/u2272744/Downloads/cyberexpertisediversity_survey_data-main/data/cyber_expertise_diversity_survey_scenarios.csv")

```

The survey was distributed on social media and was accepting responses between `r format(min(date(responses$recorded_date)), "%d/%m/%Y")` and `r format(max(date(responses$recorded_date)), "%d/%m/%Y")`

### The comb concept

-   The **personal comb** is just the values that a person put in for their own expertise using the sliders. NB for these alignment calculations we're only using those 'teeth' of the comb that correspond to cybok categories (we have some extras in but for these calculations we ignore them).

-   The **specialism comb** is the cybok values that correspond to each specialism: we create these by substituting 30 for 'wider', 60 for 'related' and 90 for 'core'.

-   The **scenario comb** is the values that a person put in for that specific scenario using the sliders. Again, for these calculations we're only using the 'teeth' that correspond to cybok categories and ignoring our extras.

```{r}
#| echo: false
#| warning: false
#personal_comb <- read.csv("../data/raw/cyber_expertise_diversity_survey_responses.csv")
```

### The research process

### The alignment metric calculation

In order to see how well (or bad) are two combs aligned, we will compute an **alignment index** $A_{(x,y)}$ as follows.

Given a set of **test values** $X:[x_0, x_1, ... x_n ]$ and a corresponding set of **reference values** $Y:[y_0, y_1, ... y_n ]$, the alignment index will be:

$$
A_{(x,y)} = \frac{\sum I(X)}{\sum Y}
$$

Where the indicator function $I(X)$ is defined as follows:

-   $I(X) = x_m$ when $x_m \leq y_m$
-   $I(X) = y_m$ when $x_m > y_m$

The indicator function is used to prevent a test comb getting 'credit' for overshooting the reference values (which could offset their deficit in other areas). To do so, it takes the reference value as a ceiling in each tooth of the test comb. This also means that the maximum possible alignment is 1.

With this, we will be calculating two types of alignments:

-   **Personal alignment**, where test values ($X$) are the person's comb (responses) and the reference values ($Y$) are the specialism comb

-   **Scenarios alignment**, where test values ($X$) are the specialism comb and the reference values ($Y$) are the scenario comb.

In both cases, the **alignment index** will result in a number between `0` and `1`, where `1` means perfect alignment between the test value and the reference value and `0` means not aligned at all.

::: callout-note
Computations for the alignment metric can be audited at the file `alignment_metric.R`
:::

## Results

The survey was conducted between `r min(responses$recorded_date)` and `r max(responses$recorded_date)` and received responses from `r nrow(responses)` individuals and describing `r nrow(scenarios)` scenarios. Of these respondents `r nrow(filter(responses, gender == 'Male'))` identified as men, `r nrow(filter(responses, sexual_orientation == 'No'))` as heterosexual, `r nrow(filter(responses, transgender == 'No'))` as their sex assigned at birth, `r nrow(filter(responses, disabilities == 'No, I do not have a disability, or a history/record of having one'))` as having no disabilities, and `r nrow(filter(responses, uk_ethnic_origin == 'White'))` of White British Ethnic origin. These figures, while not reflective of the diversity of wider British society, they are inline with previous findings (XXX ???) of the lack of diversity within the UK cyber security. 


```{r}
#| echo: false
#| warning: false

age <- ggplot(responses, aes(age)) +
          geom_bar() +
          xlab('') + 
          ylab('') +
          ggtitle('Age')

exp <- ggplot(responses, aes(cut(responses$years_experience, c(-Inf, 5, 10, 15, 20, 25, Inf), c('0-5', '6-10', '11-15', '16-20', '21-25', '25+')))) +
          geom_bar() +
          xlab('') + 
          ylab('') +
          ggtitle('Years Experience')

grid.arrange(age, exp, nrow = 1)

```


```{r}
#| echo: false
#| warning: false


# this might not be the neatest way to create this table, but I was thinking about
# how the code to create the table could be in the report without it revealing
# any of the categories we are suppressing

# First make a table of sectors, then set sector names as charecters (otherwise,
# replace will not work with factors), then set 3 or below to other, then group
# and count results

industry_table <- data.frame(table(responses$organisation_sector))
industry_table$Var1 <- industry_table$Var1 <- as.character(industry_table$Var1)
industry_table$Var1 <- replace(industry_table$Var1, industry_table$Freq <= 3, 'Other (low n)')
industry_table <- industry_table %>% group_by(Var1) %>% summarise(Freq=sum(Freq))

# order and set col names
industry_table <- industry_table %>% arrange(desc(Freq))
colnames(industry_table) <- c('Industry Sector', 'Number of respondents')

kable(industry_table)

```

```{r}
#| echo: false
#| warning: false

org_size <- data.frame(table(responses$organisation_size))

org_size <- org_size %>% arrange(factor(Var1, levels=c('Large (250 employees or more)',
                                           'Medium (50 to 249 employees)',
                                           'Small (10 to 49 employees)',
                                           'Micro (1 to 9 employees)')))

colnames(org_size) <- c('Organisation Size', 'Number of respondents')

kable(org_size)

```
### Megacombs


```{r}
#| echo: false
#| warning: false

response_all <- responses[8:33]

response_all <- pivot_longer(response_all, 
                                cols = names(response_all), 
                                values_to = 'all_points')

response_all$all_name <-  'name'
# chopping off the name variable
response_all$name <- gsub("expertise_comb_", '', response_all$name)
# 87 NAs
response_all <- drop_na(response_all)

ggplot(response_all, aes(all_points, all_name)) +
  geom_quasirandom(size= 0.5, alpha=0.5) +
  geom_boxplot(width=0.05) +
  xlab('') + 
  ylab('') +
  ggtitle('All comb responses')
```

Good distribution of responses across the sliders, with  `r 100 / nrow(response_all) * nrow(filter(response_all, all_points == 0))`% responses being set to 0 and   `r 100 / nrow(response_all) * nrow(filter(response_all, all_points == 100))`% set to 1.

::: callout-caution
## Questions for Viz

1.  Do we have a good set of descriptive views of the basic raw data? You produced quite a few, so it is just a matter of going back to these and seeing if they're ready to go.

2.  Are there features of the data that these visualisations reveal that aren't currently reflected in the report, that we should focus more on?
:::

## Limitations

### Sample size

### Sampling strategy

### Sliders
